---
title: "Stat 365: Statistical Communication"
subtitle: "Spring 2023"
# author: "Dr. Emily Robinson"
# date: "January 9, 2023"

format: 
  html:
    self-contained: true
    toc: true
  pdf:
    keep-tex: false
    # include-in-header: "in-header.tex"
    include-in-header:
      text: |
        \makeatletter
        \def\@maketitle{%
          \begin{center}%
          \let \footnote \thanks
            {\LARGE \@title \par}%
            {\large \@author \par}%
            {\large \@date}
          \end{center}%
          \par
          \vskip 1em}
        \makeatother
        \RedeclareSectionCommand[beforeskip=1ex plus -.2ex minus -.2ex,afterskip=.25ex plus -.1ex minus -.1ex]{section}
        \RedeclareSectionCommand[beforeskip=1ex plus -.2ex minus -.2ex,afterskip=.25ex plus -.1ex minus -.1ex]{subsection}
        \RedeclareSectionCommand[beforeskip=1ex plus -.2ex minus -.2ex,afterskip=.25ex plus -.1ex minus -.1ex]{subsubsection}
        \raggedbottom
        \usepackage{enumitem}
        \setlist{nolistsep}
    latex-engine: xelatex

dev: png
geometry: margin=1in

fontfamily: mathpazo
fontsize: 11pt
---

## Instructor

**Dr. Emily Robinson** *(she/her)*

-	Email: [erobin17@calpoly.edu](mailto:erobin17@calpoly.edu?subject=Stat%20331/531)
-	Office: Building 25 Office 103 (by Statistics Department Office) 

## Course information

**Class Meeting Times:** Mondays/Wednesdays

- Section 01: 1:10pm - 2:00pm
- Section 02: 2:10pm - 3:00pm

**Class Location:** 33-457 (Math & Science)

**Office Hours:** Thursdays at 2:30pm - 4:30pm, in-person (25-103)

```{r}
#| include: false
library(tidyverse)
library(lubridate)

# Create a calendar for your syllabus ----
# Source: http://svmiller.com/blog/2020/08/a-ggplot-calendar-for-your-semester/
# 1) what is the first Monday of the semester?
# Any number of ways to identify dates in R, but we'll use {lubridate} and the ymd() function here.
# Format: YYYYMMDD. In this example, 4 January 2022.

# What are the full dates of the semester?
semester_dates <- seq(ymd(20230402), ymd(20230617), by=1)
# Weekday(s) of class
class_wdays <- c("Mon", "Wed")

not_here_dates <- c(
  ymd("20230529")
  # Break
  # seq(ymd(20221017), ymd(20221018), by = 1)
  )

exam_dates <- c()
# project_dates <- c(ymd(20221104), ymd(20221205), ymd(20221209))
finals_week <- seq(ymd(20230611), ymd(20230617), by = 1)

# Custom function for treating the first day of the month as the first week 
# of the month up until the first Sunday 
# (unless Sunday was the start of the month)
wom <- function(date) {
    first <- wday(as.Date(paste(year(date), month(date), 1, sep = "-")))
    return((mday(date) + (first - 2)) %/% 7 + 1)
  }
# Create a data frame of dates, assign to Cal
Cal <- tibble(date = seq(ymd(20230401), ymd(20230620), by=1))  %>%
  mutate(mon = lubridate::month(date, label=T, abbr=F),
         wkdy = weekdays(date, abbreviate=T),
         wkdy = fct_relevel(wkdy, "Sun", "Mon", "Tue", "Wed", "Thu","Fri","Sat"), # make sure Sunday comes first
         semester = date %in% semester_dates, 
         # project = date %in% project_dates,
         exams = date %in% exam_dates, # is it an exam?
         not_here = date %in% not_here_dates, # is it a day off?
         exam_wk = date %in% finals_week,
         day = lubridate::mday(date), 
         week = wom(date))
Cal <- Cal %>%
  mutate(category = case_when(
    # project ~ "Project",
    exams ~ "Exam",
    not_here ~ "Cal Poly Holiday",
    semester & wkdy %in% class_wdays & !not_here & !exam_wk | date == "2023-01-17" ~ "Class Day",
    semester ~ "Finals",
    TRUE ~ "NA"
  )) 
class_cal <- Cal %>% 
  ggplot(.,aes(wkdy, week)) +
  theme_bw() +
  theme(aspect.ratio = 1, 
        panel.grid.major.x = element_blank(),
        legend.position = "right",
        # legend.position = c(1, 0), 
        # legend.justification = c(1,0),
        # legend.direction = "vertical", 
        legend.title = element_blank(),
        axis.title.y = element_blank(), 
        axis.title.x = element_blank(),
        axis.ticks.y = element_blank(), 
        axis.text.y = element_blank()) +
  geom_tile(alpha=0.8, aes(fill=category), color="black", size=.45) +
  facet_wrap(~mon, scales = "free", ncol=3) +
  geom_text(aes(label = day, color = semester & (!not_here))) +
  # put your y-axis down, flip it, and reverse it
  scale_y_reverse(breaks=NULL) +
  # manually fill scale colors to something you like...
  scale_color_manual(values = c("FALSE" = "grey80", "TRUE" = "black"), guide = "none") + 
  scale_fill_manual(values=c("Class Day"="purple", 
                             "Finals"="white",
                             "Cal Poly Holiday" = "grey10",
                             "NA" = "white", # I like these whited out...
                             "Exam"="orange"),
                    #... but also suppress a label for a non-class semester day
                    breaks=c("Finals", "Cal Poly Holiday", "Class Day","Exam"))
class_cal

```

# Course description

Communicating clearly is crucial for success as a statistician and for all who engage in analyzing data. The Statistics Department decided to institute this two-unit course a few years ago, in order to provide more guidance, practice, and feedback about writing well with statistics and data.

This course is required for Statistics majors but does not count toward the Statistics minor or for any other degree program.

**Prerequisites:**

## Course goals

+ Critique technical and media reports based on statistical studies and analyses.
+ Select the appropriate descriptive and inferential statistical methods (seen in previous classes), and analyze data using those methods.
+ Synthesize output from statistical software relevant to answering the research questions
+ Write reports that clearly convey statistical ideas and content; effectively incorporate appropriate
information in tables and graphs within technical reports; and integrate sufficient documentation to
ensure reproducibility.
+ Evaluate an audience's statistical background (introductory or advanced) and modify the style and
content of a written report based on the purpose, audience, and context of a given communicative
act (such as a consultation memo, executive summary, or technical report).
+ Use the following statistical concepts with lexical precision in written reports: parameter vs.
statistic, variability, p-value, margin-of-error, interaction, scope of conclusions.
+ Learn appropriate methods for visualizing and communicating data, both numerical and categorical.
+ Develop technical skills using spreadsheets, Tableau, and R, to visualize and communicate data.
+ Apply course material to communications you find in the wild, and datasets that interest you.
At the end of this class, students should be able to:
+ Use computational tools to prepare data for storytelling. This includes:
    + wrangling datasets using spreadsheets,
    + making data visualizations using DataWrapper, and
    + curating the final presentation of products.
+ Critically evaluate data stories in a field of interest. This includes:
    + describing and critiquing data visualizations and narratives including in a peer review setting,
    + evaluating structure, design, and writing choices of data stories, and
    + articulating their own data storytelling style.
+ Communicate insights from data clearly and precisely. This includes:
    + identifying a compelling story from a dataset,
    + telling the story effectively using appropriate visual and descriptive approaches,
    + knowing the audience and communicating at an accessible level that is still faithful to the data,
and communicating via a variety of mediums including written, visual, and oral products.

# Textbooks

Most textbooks and readings have been reserved through the Cal Poly Library system ([access here]()).

The main textbooks used in this course are:

+ Communicating with Data: The Art of Writing for Data Science (Deborah Nolan and Sara Stoudt)
+ Storytelling with Data (Cole Knaflic)

In addition to the textbooks, there will be a number of readings provided from other texts, including:

+ Numbers in the Newsroom (Sarah Cohen)
+ The Functional Art (Alberto Cairo)
+ Visualize This (Nathan Yau)
+ How to Lie with Statistics (Darrell Huff)
+ Show me the Numbers (Sephen Few)
+ The Visual Display of Quantitative Information (Edward Tufte)
+ Better Data Visualizations (Jonathan Schwabish)

and more.

# Grading

In this class we will be using a contract grading system. This is designed to give flexibility and freedom to explore while ensuring a level of accountability. Learning to write takes practice. I do not want worries
about grades to distract us from being creative, taking risks, and ultimately finding our voices.

You are guaranteed a grade of a B for this class if you meet the following conditions:

+ Complete readings, readiness quizzes, writing prompts, and class activities on time, following the instructions meeting "satisfactory", and missing no more than 3.
+ Complete each mini-project on time, following the instructions and meeting the conditions of "satisfactory" on the rubric.
+ Complete the final report on time, following the instructions and meet the conditions of "satisfactory" on the rubric.

Your grade will decrease by a third of a step (for example a B to a B-) for each B condition that you do not meet. Missing assignments will be scaled, so that if for example you miss 6 ,
turn in one mini-project late, or turn in two projects late your grade will decrease by two thirds of a step (for example
a B to a C+).

You can increase your letter grade by a third of a step (for example a B to B+) for meeting each of the following conditions (further instructions will be provided):

+ Complete a weekly submission to the [NYT: What's going on in this graph](https://www.nytimes.com/column/whats-going-on-in-this-graph), missing no more than two weeks throughout the quarter.
+ Select a data set from [TidyTuesday](https://github.com/rfordatascience/tidytuesday) or [Makeover Monday](https://www.makeovermonday.co.uk/data/) and use guiding principles to create an effective dashboard using Tableau or RShiny to highlight key points.
+ [Make your own website](https://earobinson95.github.io/stat365-calpoly/98-A-options/make-your-own-website.html) using Quarto and GitHub to showcase your statistical projects.
+ Complete a final project that goes above and beyond the requirements. Your job is to convince me in an additional one page document why you think your contribution goes the extra mile.

For example, to earn an A in this class you must meet the conditions for a B and complete three of the "boosting" conditions above.

Any deviation from the grading policies outlined above will only be to your benefit.

## Grade Breakdown

1. **Readings, readiness quizzes, and participation**

2. **Writing prompts, and activities**

3. **Mini projects:** There will be four mini-projects throughout the semester. The mini projects include Dear Data, One Number Story, Copy the Masters, and Build Your Portfolio. Mini-project portions will be graded on 4 point scales.

4. **Final project:**

## Late work


# Class Schedule

This schedule is tentative and subject to change.

```{r calendar}
#| echo: false
#| eval: true
#| warning: false
#| fig-width: 8
#| fig-height: 3.5
#| fig-pos: 'h'
class_cal
```

```{r schedule}
#| echo: false
#| eval: true
#| fig-pos: 'h'
# class_schedule |> 
# kableExtra::kable(caption = "Tentative schedule of class topics and important due dates", format = "simple")
```

# Course Expectations

You will get out of this course what you put in. The following excerpt was taken
from Rob Jenkins’ article “Defining the Relationship” which was published in The
Chronicle of Higher Education (August 8, 2016). This accurately summarizes what
I expect of you in my classroom (and also what you should expect of me).

*“I’d like to be your partner. More than anything, I’d like for us to form a mutually beneficial alliance in this endeavor we call education.*

*I pledge to do my part. I will:*

+ *Stay abreast of the latest ideas in my field.* 
+ *Teach you what I believe you need to know; with all the enthusiasm I possess.*
+ *Invite your comments and questions and respond constructively.*
+ *Make myself available to you outside of class (within reason).*
+ *Evaluate your work carefully and return it promptly with feedback.* 
+ *Be as fair, respectful, and understanding as I can humanly be.*
+ *If you need help beyond the scope of this course, I will do my best to provide it or see that you get it.* 

*In return, I expect you to:*

+ *Show up for class each day or let me know (preferably in advance) if you have some good reason to be absent.* 
+ *Do your reading and other assignments outside of class and be prepared for each class meeting.*
+ *Focus during class on the work we’re doing and not on extraneous matters (like whoever or whatever is on your phone at the moment).* 
+ *Participate in class discussions.*
+ *Be respectful of your fellow students and their points of view.* 
+ *In short, I expect you to devote as much effort to learning as I devote to teaching.*

*What you get out of this relationship is that you’ll be better equipped to succeed in this and other college courses, work-related assignments, and life in general. What I get is a great deal of professional and personal satisfaction. Because I do really like you [all] and want the best for you.”*

# University Policies

See [academicprograms.calpoly.edu/content/academicpolicies](https://academicprograms.calpoly.edu/content/academicpolicies/index).

## Learning Environment and Support

It is my goal for everyone to feel safe and comfortable in my classroom. If there is any way I can make the course more welcoming for you, please do not hesitate to ask.

In particular, if you have a disability, I will gladly work with you to make this class accessible.

I encourage you to also contact the Disability Resource Center (Building 124, Room 119 or at 805-756-1395), who can help you register for extra accommodations such as extended exam time.

If you are having difficulty affording groceries, lacking a safe & stable place to live, or needing additional essential supports, please see Canvas for a list of Student Support Services at Cal Poly.

## Academic Integrity and Class Conduct

Simply put, I will not tolerate cheating or plagiarism.

Any incident of dishonesty, copying, or plagiarism will be
reported to the Office of Student Rights and Responsibilities.

Cheating will earn you an incomplete on the assignment and an overall 1/3 grade drop penalty (e.g. B to a B-). In circumstances of flagrant cheating, you may be given a grade of F in the course.

Paraphrasing or quoting another’s work without citing the source is a form of academic misconduct.

For more information about what constitutes cheating and plagiarism, please see
[academicprograms.calpoly.edu/content/academicpolicies/Cheating](https://academicprograms.calpoly.edu/content/academicpolicies/Cheating).

