---
title: "Stat 365: Statistical Communication"
subtitle: "Spring 2024"
# author: "Dr. Emily Robinson"
# date: "January 9, 2023"
format: 
  html:
    self-contained: true
    toc: true
    theme:
      light: flatly
      dark: darkly
dev: png
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
editor: 
  markdown: 
    wrap: 80
---

## Instructor

**Dr. Emily Robinson** *(she/her)*

-   Email:
    [erobin17\@calpoly.edu](mailto:erobin17@calpoly.edu?subject=Stat%20365)
-   Office: Building 25 Office 103 (by Statistics Department Office)

## Course information

**Class Meeting Times:** Mondays/Wednesdays

-   Section 01: 8:10pm - 9:00pm
-   Section 02: 9:10pm - 10:00pm

**Class Location:** 010 - 0227

**Office Hours:** in-person (25-103)

- Wednesdays at 1:10pm - 2:30pm
- Thursdays at 9:10 - 10:30am

```{r}
#| include: false
library(tidyverse)

# Create a calendar for your syllabus ----
# Source: http://svmiller.com/blog/2020/08/a-ggplot-calendar-for-your-semester/
# 1) what is the first Monday of the semester?
# Any number of ways to identify dates in R, but we'll use {lubridate} and the ymd() function here.
# Format: YYYYMMDD. In this example, 4 January 2022.

# What are the full dates of the semester?
semester_dates <- seq(ymd(20240401), ymd(20240608), by = 1)
# Weekday(s) of class
class_wdays <- c("Mon", "Wed")

not_here_dates <- c(
  ymd("20240527"),
  ymd("20240401")
  # Break
  # seq(ymd(20221017), ymd(20221018), by = 1)
  )

exam_dates <- c(ymd("20240610"),
                ymd("20240612")
                )
# project_dates <- c(ymd(20221104), ymd(20221205), ymd(20221209))
finals_week <- seq(ymd(20240610), ymd(20240615), by = 1)

# Custom function for treating the first day of the month as the first week 
# of the month up until the first Sunday 
# (unless Sunday was the start of the month)
wom <- function(date) {
    first <- wday(as.Date(paste(year(date), month(date), 1, sep = "-")))
    return((mday(date) + (first - 2)) %/% 7 + 1)
  }
# Create a data frame of dates, assign to Cal
Cal <- tibble(date = seq(ymd(20240401), ymd(20240630), by = 1))  %>%
  mutate(mon = lubridate::month(date, label=T, abbr=F),
         wkdy = weekdays(date, abbreviate=T),
         wkdy = fct_relevel(wkdy, "Sun", "Mon", "Tue", "Wed", "Thu","Fri","Sat"), # make sure Sunday comes first
         semester = date %in% semester_dates, 
         # project = date %in% project_dates,
         exams = date %in% exam_dates, # is it an exam?
         not_here = date %in% not_here_dates, # is it a day off?
         exam_wk = date %in% finals_week,
         day = lubridate::mday(date), 
         week = wom(date)
         )

Cal <- Cal %>%
  mutate(category = case_when(
    # project ~ "Project",
    exams ~ "Final Slots",
    not_here ~ "No Class",
    semester & wkdy %in% class_wdays & !not_here & !exam_wk | date == "2024-05-28" ~ "Class Day",
    semester ~ "Finals",
    TRUE ~ "NA"
  )) 
class_cal <- Cal %>% 
  ggplot(.,aes(wkdy, week)) +
  theme_bw() +
  theme(aspect.ratio = 1, 
        panel.grid.major.x = element_blank(),
        legend.position = "right",
        # legend.position = c(1, 0), 
        # legend.justification = c(1,0),
        # legend.direction = "vertical", 
        legend.title = element_blank(),
        axis.title.y = element_blank(), 
        axis.title.x = element_blank(),
        axis.ticks.y = element_blank(), 
        axis.text.y = element_blank()) +
  geom_tile(alpha=0.8, aes(fill=category), color="black", size=.45) +
  facet_wrap(~mon, scales = "free", ncol=3) +
  geom_text(aes(label = day, color = semester & (!not_here))) +
  # put your y-axis down, flip it, and reverse it
  scale_y_reverse(breaks=NULL) +
  # manually fill scale colors to something you like...
  scale_color_manual(values = c("FALSE" = "grey80", "TRUE" = "black"), guide = "none") + 
  scale_fill_manual(values=c("Class Day"="purple", 
                             "Finals"="white",
                             "No Class" = "grey10",
                             "NA" = "white", # I like these whited out...
                             "Final Slots"="orange"),
                    #... but also suppress a label for a non-class semester day
                    breaks=c("Finals", "No Class", "Class Day", "Final Slots"))
class_cal
```

# Course description

This two-unit course is designed to help students develop strong written
communication skills in statistics and data analysis. Through guided practice
and feedback, students will learn to effectively communicate statistical ideas
and results to different audiences using appropriate graphs and tables. The
course will cover strategies for discerning relevant and necessary information
and for selecting and describing appropriate methods to analyze data, building
on students' previous statistics coursework.

**Prerequisites:** Completion of GE Area A2, completion of GE Area A3, and a
second course in Statistical Methods (STAT 252, STAT 302, or STAT 313).

**Requirements:** This course is required for Statistics majors but does not
count toward the Statistics minor or any other degree program. Please speak to your academic adviser
if you need permission for credit.

## Course goals

By the end of this course, you will:

-   Enhance your written communication skills in statistics and data analysis,
    specifically through the use of appropriate graphs and tables to convey
    statistical ideas and results.
-   Synthesize output from statistical software to address research questions
    and analyze data with lexical precision and accuracy.
-   Evaluate the statistical background of your intended audience, and adapt the
    style and content of your written reports to suit their needs and context.
-   Demonstrate your understanding of statistical concepts such as parameters
    versus statistics, variability, p-values, margin-of-error, interaction, and
    scope of conclusions.
-   Develop proficiency in writing and producing technical reports with adequate
    documentation to ensure reproducibility.
-   Learn how to use spreadsheets and R for data visualization and
    communication.
-   Gain experience in integrating statistical analysis, data visualization, and
    technical writing in a reproducible manner.

# Textbooks

I have reserved eBooks for most of the textbooks and readings through the Cal
Poly Library [course reserves](https://reserves.calstate.edu/slo/).

The main textbooks used in this course are:

-   Communicating with Data: The Art of Writing for Data Science (Deborah Nolan
    and Sara Stoudt)
-   Storytelling with Data (Cole Knaflic)

In addition to the textbooks, there will be a number of readings provided from
other texts, including:

-   Numbers in the Newsroom (Sarah Cohen)
-   The Functional Art (Alberto Cairo)
-   Visualize This (Nathan Yau)
-   How to Lie with Statistics (Darrell Huff)
-   Show Me the Numbers (Stephen Few)
-   The Visual Display of Quantitative Information (Edward Tufte)
-   Better Data Visualizations (Jonathan Schwabish)

and more.

# Grading

In this class we will be using a contract grading system. This is designed to
give flexibility and freedom to explore while ensuring a level of
accountability. Learning to write takes practice. I do not want worries about
grades to distract us from being creative, taking risks, and ultimately finding
our voices.

You are guaranteed a grade of a B for this class if you meet the following
conditions:

-   Submit assignments (readings, class activities, peer
    reviews, and final group report check-points) on time and meet the "satisfactory" criteria with a maximum of 3 missed assignments.
-   Complete all mini-projects on time and meet the "satisfactory" conditions on
    the rubrics.
-   Submit the final group report on time and meet the "satisfactory"
    conditions on the rubric.

Your grade will decrease by a third of a step (for example a B to a B-) for each
B condition that you do not meet. Missing assignments will be scaled, so that if
for example you miss six assignments, turn in two mini-projects late, or turn in
your final group report late your grade will decrease by two thirds of a
step (for example a B to a C+).

A missing final technical report submission will result in a maximum grade of a
D+ at the instructor's discretion.

You can increase your letter grade by *up to* a third of a step (for example a B
to B+) for meeting each of the following conditions to satisfactory as defined on the rubrics (further instructions will
be provided):

-   Complete a weekly submission to the [NYT: What's going on in this
    graph](https://www.nytimes.com/column/whats-going-on-in-this-graph), missing
    no more than one week throughout the quarter.
-   Start a [Dear Data]() "pen pal" with a peer, including at least 3 thoughtful entries.
-   Listen to the [99% Invisible podcast on Florence Nightingale: Data Viz
    Pioneer](https://99percentinvisible.org/episode/florence-nightingale-data-viz-pioneer/). Find the data set to recreate the rose diagram, compare with a stacked bar plot, discuss the story, and argue why Florence needed the rose diagram.
-   [Copy the Masters]() by recreating a graphic from FiveThirtyEight using your favorite software.
-   [Make your own
    website](https://earobinson95.github.io/stat365-calpoly/98-grade-boosters/make-your-own-website.html) using Quarto and GitHub pages to showcase your statistical projects.
-   In addition to the mini-project for building your Resume, find an internship/program and write a [Tailored Cover Letter]().
-   Complete a final group report project that goes above and beyond the
    requirements. Your job is to convince me in an additional half page document
    why you think your contribution goes the extra mile.
-   I encourage you to suggest your own "boosting" condition via the suggestion
    "box" form on Canvas.

For example, to earn an A in this class you must meet the conditions for a B and
complete three of the "boosting" conditions above to "satisfactory" quality.

Any deviation from the grading policies outlined above will only be to your
benefit.

## Grade Breakdown

1.  **Readings and participation:**

You will be expected to complete assigned readings before coming to class and
participate in class discussions and activities. Your peer review participation
and engagement in class will be evaluated throughout the quarter.
<!-- Additionally, you will be required to take readiness quizzes to assess your understanding of the material. -->

2.  **In-class activities:**

Most classes will involve activities to help you develop your technical writing and communication skills. These will include assignments such as summarizing statistical findings, redesigning graphs and tables, and story-boarding your data analysis. These assignments will be designed to help you practice communicating complex statistical concepts and analyses clearly and effectively.

3.  **Mini-projects:**

There will be three mini-projects throughout the quarter, each designed to help
you develop and apply your statistical communication skills. Additionally, the first two mini-projects will build into your final group report. Mini-projects are to be completed *individually.* The mini-projects will include:

-   *One Number Story:* In this assignment, you will choose a single number that
    summarizes a key finding from the data set your group selected for your final report and craft a compelling story around it, using appropriate statistical evidence to support your claims.

-   *Stand Alone Graphic:* For this assignment, you will xxxx

-   *Build your resume:* In this assignment, you will create a resume using
    Overleaf to highlight your education, statistical communication and
    technical skills, statistics and data science projects, and other relevant
    experiences.

Detailed rubrics will be provided for each mini-project to help guide your work
and ensure that you meet the required standards.

4.  **Group final report:**

The group final report is the culminating assignment of the course, where you will have the opportunity to apply all the skills and knowledge you have acquired throughout the quarter.
This will be completed in *groups of three to four individuals*. You will be required to choose a data set from [TidyTuesday](https://github.com/rfordatascience/tidytuesday) and investigate a research question related to the data. You will analyze the data using appropriate statistical methods and
produce a technical report that effectively communicates your findings in written, graphical, and table format. You will
be provided with guidelines and resources to support your work, and you will
receive feedback and guidance throughout the project. The group report will be evaluated based on its clarity, thoroughness, and adherence to standard statistical practices.

A portion of the rubric will be dedicated to effective and equitable collaboration efforts.

## Late work

I understand that unexpected events may arise during the quarter that may
prevent you from submitting your assignments on time. To accommodate for this,
you will be given three 24-hour late submission "Grace Days" to use on In-class Activities and Mini-projects throughout the quarter. To use a "Grace Day", you must (1) send me an email and (2) comment on the Canvas submission.

These "Grace Days" **cannot** be used on your final group report submission.

Your remaining "Grace Days" will be tracked in a Canvas assignment as three points, with each "Grace Day" worth one point.

## Live Revisions

In-class Activities and Mini-projects are subject to revisions if "satisfactory" is not met. If you do not receive, satisfactory, you are expected to revise your submission based on my feedback. For In-class activities, please stop by my office hours to talk me through your changes and why it now meets satisfactory requirements. For mini-projects, please schedule a 10-minute meeting with me via a provided Calendly link on Canvas (this will be set up for specific weeks after mini-project due dates).

Booster Conditions will be given "Best By" dates for deadlines. If you turn the assignment in by the "Best By" date, you will receive feedback and an opportunity to revise before the end of the quarter. Booster conditions may still be accepted past the "Best By" date, but will not be graded until finals week leaving no opportunity for revisions if the instructor determines "satisfactory" quality is not met.

*Disclaimer: I've never done Live Revisions, but wanted to try it!"*

# Class Schedule

```{r calendar}
#| echo: false
#| eval: true
#| warning: false
#| fig-width: 8
#| fig-height: 3.5
#| fig-pos: 'h'
#| fig-cap: "Note: Tuesday, May 28th follows a Monday Schedule."
class_cal
```

This schedule is **very** tentative and subject to change.

+-------------------------+-------------------------+-------------------------+------------------+
| Week                    | Topics                  | Readings                | Main Deliverables|
+=========================+=========================+=========================+==================+
| 1: 4/3                  | Writing about Data      | CwD (ch. 1, ch. 10.3),  |                  |
|                         |                         | How to Lie with Stats   |                  |
|                         |                         | (ch. 10)                |                  |
+-------------------------+-------------------------+-------------------------+------------------+
| 2: 4/8, 4/10            | Final Project +         | CwD (ch. 3.1 - 3.3)     | Initial Data     |
|                         | Reproducibility         |                         | Proposals        |
|                         |                         |                         |                  |
|                         | Data Descriptions       |                         |                  |
|                         |                         |                         |                  |
+-------------------------+-------------------------+-------------------------+------------------+
| 3: 4/15, 4/17           | Data Journalism         | CwD (ch. 2), Numbers in | One Number Story |
|                         |                         | the Newsroom (ch. 1, 2) |                  |
|                         | Spreadsheets, Summaries,|                         |                  |
|                         | and Pivot Tables        |                         |                  |
+-------------------------+-------------------------+-------------------------+------------------+
| 4: 4/22, 4/24           | Peer Review One-Number  |                         |                  |
|                         | stories                 |                         |                  |
|                         |                         |                         |                  |
|                         | Table Design            |                         |                  |
|                         |                         |                         |                  |
+-------------------------+-------------------------+-------------------------+------------------+
| 5: 4/29, 5/1            | Intro/History of Viz    | Storytelling with Data  | Stand Alone      |
|                         |                         | Podcast Episodes #27:   | Graphic          |
|                         |                         | what is data            |                  |
|                         | Accessibility, Colors,  | visualization and #67:  |                  |
|                         | and Perception          | beginner mistakes in    |                  |
|                         |                         | data viz                |                  |
|                         |                         |                         |                  |
|                         |                         | Storytelling with Data  |                  |
|                         |                         | Episode #8: The many    |                  |
|                         |                         | myths of data           |                  |
|                         |                         | visualization           |                  |
|                         |                         |                         |                  |
|                         |                         | (Optional) Show me the  |                  |
|                         |                         | numbers (ch. 5)         |                  |
+-------------------------+-------------------------+-------------------------+------------------+
| 5: 5/6, 5/8             | Data Viz Makeovers      |                         |                  |
|                         |                         |                         |                  |
|                         | How to Read a Research  | CwD ch. 7.1 - 7.5       |                  |
|                         | Paper                   | CwD ch. 8.1 - 8.7       |                  |
|                         |                         |                         |                  |
+-------------------------+-------------------------+-------------------------+------------------+
| 6: 5/13, 5/15           | Storyboarding           | CwD ch. xxxx            | Group Storyboard |
|                         |                         |                         |                  |
|                         |                         |                         |                  |
+-------------------------+-------------------------+-------------------------+------------------+
| 8: 5/20, 5/22           | Visit from Career       |                         | Build Your Resume|
|                         | Center                  |                         |                  |
|                         |                         |                         |                  |
|                         | Finding Internships +   |                         |                  |
|                         | Online Presence         |                         |                  |
+-------------------------+-------------------------+-------------------------+------------------+
| 9: 5/28, 5/29           | Understanding the       |                         | Final Report     |
|                         | Publication Process     |                         | Draft            |
|                         |                         |                         |                  |
|                         | Project Workdays        |                         |                  |
|                         |                         |                         |                  |
+-------------------------+-------------------------+-------------------------+------------------+
| 10: 6/3, 6/5            | Project Revisions       |                         |                  |
|                         |                         |                         |                  |
|                         | **No Class Wednesday**  |                         |                  |
+-------------------------+-------------------------+-------------------------+------------------+
| Finals                  | **Mon., June 10th       |                         | Final Report     |
|                         | at** **7:10 - 10am:**   |                         | Revisions        |
| *Reserve these times    | Section 01 *(if you     |                         |                  |
| for potential final     | come to class 8:10 -    |                         |                  |
| grade meetings.*        | 9am)*                   |                         |                  |
|                         |                         |                         |                  |
|                         | **Wed., June 12th at    |                         |                  |
|                         | 7:10 - 10am:** Section  |                         |                  |
|                         | 02 *(if you come to     |                         |                  |
|                         | class 9:10 - 10am)*     |                         |                  |
+-------------------------+-------------------------+-------------------------+------------------+

# Course Expectations

You will get out of this course what you put in. The following excerpt was taken
from Rob Jenkins' article "Defining the Relationship" which was published in The
Chronicle of Higher Education (August 8, 2016). This accurately summarizes what
I expect of you in my classroom (and also what you should expect of me).

*"I'd like to be your partner. More than anything, I'd like for us to form a
mutually beneficial alliance in this endeavor we call education.*

*I pledge to do my part. I will:*

-   *Stay abreast of the latest ideas in my field.*
-   *Teach you what I believe you need to know; with all the enthusiasm I
    possess.*
-   *Invite your comments and questions and respond constructively.*
-   *Make myself available to you outside of class (within reason).*
-   *Evaluate your work carefully and return it promptly with feedback.*
-   *Be as fair, respectful, and understanding as I can humanly be.*
-   *If you need help beyond the scope of this course, I will do my best to
    provide it or see that you get it.*

*In return, I expect you to:*

-   *Show up for class each day or let me know (preferably in advance) if you
    have some good reason to be absent.*
-   *Do your reading and other assignments outside of class and be prepared for
    each class meeting.*
-   *Focus during class on the work we're doing and not on extraneous matters
    (like whoever or whatever is on your phone at the moment).*
-   *Participate in class discussions.*
-   *Be respectful of your fellow students and their points of view.*
-   *In short, I expect you to devote as much effort to learning as I devote to
    teaching.*

*What you get out of this relationship is that you'll be better equipped to
succeed in this and other college courses, work-related assignments, and life in
general. What I get is a great deal of professional and personal satisfaction.
Because I do really like you \[all\] and want the best for you."*

# Learning Environment and Support

I am committed to creating a safe and inclusive learning environment where all
students feel respected and supported. If there are any ways I can improve the
classroom environment to make it more welcoming for you, please don't hesitate
to let me know.

If you have a disability and require accommodations to fully participate in the
course, please contact me as soon as possible to discuss how I can best support
you. I also encourage you to register with Cal Poly's Disability Resource Center
(Building 124, Room 119 or at 805-756-1395) to explore additional accommodations
that may be available to you.

If you are experiencing food insecurity, housing instability, or other
challenges that may impact your ability to succeed in this course, please refer
to the resources listed on Canvas under "Student Support Services at Cal Poly."
These resources provide a range of essential support services, including
emergency financial assistance, counseling, and academic support.

I am committed to working with you to ensure that you have the resources and
support you need to succeed in this course. Let's work together to create a
positive and inclusive learning environment for all students.

# Academic Integrity

Academic integrity is a fundamental value of this course and of the university.
Simply put, I will not tolerate cheating, plagiarism, or any other form of
academic dishonesty.

Any incident of academic misconduct, including dishonesty, copying, or
plagiarism, will be reported to the Office of Student Rights and
Responsibilities.

Cheating or plagiarism will result in an incomplete grade for the assignment and
an overall grade deduction of one-third (e.g., B to a B-). In cases of flagrant
cheating, a grade of F for the course may be assigned.

It is important to note that paraphrasing or quoting another's work without
proper citation is a form of academic misconduct. This includes using Chat GPT,
which should only be used to generate ideas and not as a substitute for your own
work.

To ensure academic integrity, please be sure to cite all sources appropriately
and only use Chat GPT in an ethical manner. For more information on academic
misconduct and what constitutes cheating and plagiarism, please see
[academicprograms.calpoly.edu/content/academicpolicies/Cheating](https://academicprograms.calpoly.edu/content/academicpolicies/Cheating).

# Acknowledgments

I would like to acknowledge the contributions of various individuals whose work
has been incorporated into this course. Special thanks to Beth Chance, Allan
Rossman, Amelia McNamara, and Sara Stoudt for their valuable materials and
inspiration.
