---
title: "Stat 365: Statistical Communication"
subtitle: "Spring 2023"
# author: "Dr. Emily Robinson"
# date: "January 9, 2023"
format: 
  html:
    self-contained: true
    toc: true
    theme:
      light: flatly
      dark: darkly
  # pdf:
  #       \makeatother
  #       \RedeclareSectionCommand[beforeskip=1ex plus -.2ex minus -.2ex,afterskip=.25ex plus -.1ex minus -.1ex]{section}
  #       \RedeclareSectionCommand[beforeskip=1ex plus -.2ex minus -.2ex,afterskip=.25ex plus -.1ex minus -.1ex]{subsection}
  #       \RedeclareSectionCommand[beforeskip=1ex plus -.2ex minus -.2ex,afterskip=.25ex plus -.1ex minus -.1ex]{subsubsection}
  #       \raggedbottom
  #       \usepackage{enumitem}
  #       \setlist{nolistsep}
  #   keep-tex: false
  #   # include-in-header: "in-header.tex"
  #   include-in-header:
  #         \end{center}%
  #         \par
  #         \vskip 1em}
  #     text: |
  #       \makeatletter
  #       \def\@maketitle{%
  #         \begin{center}%
  #         \let \footnote \thanks
  #           {\LARGE \@title \par}%
  #           {\large \@author \par}%
  #           {\large \@date}
  #   latex-engine: xelatex
dev: png
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
editor: 
  markdown: 
    wrap: 80
---

## Instructor

**Dr. Emily Robinson** *(she/her)*

-   Email:
    [erobin17@calpoly.edu](mailto:erobin17@calpoly.edu?subject=Stat%20365)
-   Office: Building 25 Office 103 (by Statistics Department Office)

## Course information

**Class Meeting Times:** Mondays/Wednesdays

-   Section 01: 1:10pm - 2:00pm
-   Section 02: 2:10pm - 3:00pm

**Class Location:** 33-457 (Fisher)

**Office Hours:** Thursdays at 2:30pm - 4:30pm, in-person (25-103)

```{r}
#| include: false
library(tidyverse)
library(lubridate)

# Create a calendar for your syllabus ----
# Source: http://svmiller.com/blog/2020/08/a-ggplot-calendar-for-your-semester/
# 1) what is the first Monday of the semester?
# Any number of ways to identify dates in R, but we'll use {lubridate} and the ymd() function here.
# Format: YYYYMMDD. In this example, 4 January 2022.

# What are the full dates of the semester?
semester_dates <- seq(ymd(20230402), ymd(20230617), by=1)
# Weekday(s) of class
class_wdays <- c("Mon", "Wed")

not_here_dates <- c(
  ymd("20230529")
  # Break
  # seq(ymd(20221017), ymd(20221018), by = 1)
  )

exam_dates <- c()
# project_dates <- c(ymd(20221104), ymd(20221205), ymd(20221209))
finals_week <- seq(ymd(20230611), ymd(20230617), by = 1)

# Custom function for treating the first day of the month as the first week 
# of the month up until the first Sunday 
# (unless Sunday was the start of the month)
wom <- function(date) {
    first <- wday(as.Date(paste(year(date), month(date), 1, sep = "-")))
    return((mday(date) + (first - 2)) %/% 7 + 1)
  }
# Create a data frame of dates, assign to Cal
Cal <- tibble(date = seq(ymd(20230401), ymd(20230620), by=1))  %>%
  mutate(mon = lubridate::month(date, label=T, abbr=F),
         wkdy = weekdays(date, abbreviate=T),
         wkdy = fct_relevel(wkdy, "Sun", "Mon", "Tue", "Wed", "Thu","Fri","Sat"), # make sure Sunday comes first
         semester = date %in% semester_dates, 
         # project = date %in% project_dates,
         exams = date %in% exam_dates, # is it an exam?
         not_here = date %in% not_here_dates, # is it a day off?
         exam_wk = date %in% finals_week,
         day = lubridate::mday(date), 
         week = wom(date))
Cal <- Cal %>%
  mutate(category = case_when(
    # project ~ "Project",
    exams ~ "Exam",
    not_here ~ "Cal Poly Holiday",
    semester & wkdy %in% class_wdays & !not_here & !exam_wk | date == "2023-01-17" ~ "Class Day",
    semester ~ "Finals",
    TRUE ~ "NA"
  )) 
class_cal <- Cal %>% 
  ggplot(.,aes(wkdy, week)) +
  theme_bw() +
  theme(aspect.ratio = 1, 
        panel.grid.major.x = element_blank(),
        legend.position = "right",
        # legend.position = c(1, 0), 
        # legend.justification = c(1,0),
        # legend.direction = "vertical", 
        legend.title = element_blank(),
        axis.title.y = element_blank(), 
        axis.title.x = element_blank(),
        axis.ticks.y = element_blank(), 
        axis.text.y = element_blank()) +
  geom_tile(alpha=0.8, aes(fill=category), color="black", size=.45) +
  facet_wrap(~mon, scales = "free", ncol=3) +
  geom_text(aes(label = day, color = semester & (!not_here))) +
  # put your y-axis down, flip it, and reverse it
  scale_y_reverse(breaks=NULL) +
  # manually fill scale colors to something you like...
  scale_color_manual(values = c("FALSE" = "grey80", "TRUE" = "black"), guide = "none") + 
  scale_fill_manual(values=c("Class Day"="purple", 
                             "Finals"="white",
                             "Cal Poly Holiday" = "grey10",
                             "NA" = "white", # I like these whited out...
                             "Exam"="orange"),
                    #... but also suppress a label for a non-class semester day
                    breaks=c("Finals", "Cal Poly Holiday", "Class Day","Exam"))
class_cal

```

# Course description

This two-unit course is designed to help students develop strong written
communication skills in statistics and data analysis. Through guided practice
and feedback, students will learn to effectively communicate statistical ideas
and results to different audiences using appropriate graphs and tables. The
course will cover strategies for discerning relevant and necessary information
and for selecting and describing appropriate methods to analyze data, building on students'
previous statistics coursework.

**Prerequisites:** Completion of GE Area A2, completion of GE Area A3, and a
second course in Statistical Methods (STAT 252, STAT 302, or STAT 313).

**Requirements:** This course is required for Statistics majors but does not
count toward the Statistics minor or any other degree program.

## Course goals

By the end of this course, you will:

-   Enhance your written communication skills in statistics and data analysis,
    specifically through the use of appropriate graphs and tables to convey
    statistical ideas and results.
-   Synthesize output from statistical software to address research questions
    and analyze data with lexical precision and accuracy.
-   Evaluate the statistical background of your intended audience, and adapt the
    style and content of your written reports to suit their needs and context.
-   Demonstrate your understanding of statistical concepts such as parameters
    versus statistics, variability, p-values, margin-of-error, interaction, and
    scope of conclusions.
-   Develop proficiency in writing and producing technical reports
    with adequate documentation to ensure reproducibility.
-   Learn how to use spreadsheets, Tableau, and R for data visualization and
    communication.
-   Gain experience in integrating statistical analysis, data visualization, and
    technical writing in a reproducible manner.

<!-- -   Develop effective written communication skills in statistics and data -->

<!--     analysis. -->

<!-- -   Synthesize output from statistical software relevant to answering the -->

<!--     research questions -->

<!-- -   Learn to use appropriate graphs and tables to communicate statistical ideas -->

<!--     and results. -->

<!-- -   Evaluate an audience's statistical background (introductory or advanced) and -->

<!--     modify the style and content of a written report based on the purpose, -->

<!--     audience, and context of a given communicative act. -->

<!-- -   Use the following statistical concepts with lexical precision in written -->

<!--     reports: parameter vs. statistic, variability, p-value, margin-of-error, -->

<!--     interaction, scope of conclusions. -->

<!-- -   Gain practice in writing technical reports. -->

<!-- -   Integrate sufficient documentation to ensure reproducibility. -->

<!-- -   Develop technical skills using spreadsheets, Tableau, and R, to visualize -->

<!--     and communicate data. -->

<!-- + Critique technical and media reports based on statistical studies and analyses. -->

<!-- + Select the appropriate descriptive and inferential statistical methods (seen in previous classes), and analyze data using those methods. -->

<!-- + Synthesize output from statistical software relevant to answering the research questions -->

<!-- + Write reports that clearly convey statistical ideas and content; effectively incorporate appropriate -->

<!-- information in tables and graphs within technical reports; and integrate sufficient documentation to -->

<!-- ensure reproducibility. -->

<!-- + Evaluate an audience's statistical background (introductory or advanced) and modify the style and -->

<!-- content of a written report based on the purpose, audience, and context of a given communicative -->

<!-- act (such as a consultation memo, executive summary, or technical report). -->

<!-- + Use the following statistical concepts with lexical precision in written reports: parameter vs. -->

<!-- statistic, variability, p-value, margin-of-error, interaction, scope of conclusions. -->

<!-- + Learn appropriate methods for visualizing and communicating data, both numerical and categorical. -->

<!-- + Develop technical skills using spreadsheets, Tableau, and R, to visualize and communicate data. -->

<!-- + Apply course material to communications you find in the wild, and datasets that interest you. -->

<!-- At the end of this class, students should be able to: -->

<!-- + Use computational tools to prepare data for storytelling. This includes: -->

<!--     + wrangling datasets using spreadsheets, -->

<!--     + making data visualizations using DataWrapper, and -->

<!--     + curating the final presentation of products. -->

<!-- + Critically evaluate data stories in a field of interest. This includes: -->

<!--     + describing and critiquing data visualizations and narratives including in a peer review setting, -->

<!--     + evaluating structure, design, and writing choices of data stories, and -->

<!--     + articulating their own data storytelling style. -->

<!-- + Communicate insights from data clearly and precisely. This includes: -->

<!--     + identifying a compelling story from a dataset, -->

<!--     + telling the story effectively using appropriate visual and descriptive approaches, -->

<!--     + knowing the audience and communicating at an accessible level that is still faithful to the data, -->

<!-- and communicating via a variety of mediums including written, visual, and oral products. -->

# Textbooks

I have reserved eBooks for most of the textbooks and readings through the Cal Poly Library
[course reserves](https://reserves.calstate.edu/slo/).

The main textbooks used in this course are:

-   Communicating with Data: The Art of Writing for Data Science (Deborah Nolan
    and Sara Stoudt)
-   Storytelling with Data (Cole Knaflic)

In addition to the textbooks, there will be a number of readings provided from
other texts, including:

-   Numbers in the Newsroom (Sarah Cohen)
-   The Functional Art (Alberto Cairo)
-   Visualize This (Nathan Yau)
-   How to Lie with Statistics (Darrell Huff)
-   Show Me the Numbers (Stephen Few)
-   The Visual Display of Quantitative Information (Edward Tufte)
-   Better Data Visualizations (Jonathan Schwabish)

and more.

# Grading

In this class we will be using a contract grading system. This is designed to
give flexibility and freedom to explore while ensuring a level of
accountability. Learning to write takes practice. I do not want worries about
grades to distract us from being creative, taking risks, and ultimately finding
our voices.

You are guaranteed a grade of a B for this class if you meet the following
conditions:

- Submit assignments (readings, readiness quizzes, writing prompts, class activities, peer reviews, and final report check-points) on time and meet the "satisfactory" criteria with a maximum of 3 missed assignments.
- Complete all mini-projects on time and meet the "satisfactory" conditions on the rubrics.
- Submit the final technical report on time and meet the "satisfactory" conditions on the rubric.

Your grade will decrease by a third of a step (for example a B to a B-) for each
B condition that you do not meet. Missing assignments will be scaled, so that if
for example you miss six assignments, turn in two mini-projects late, or turn in
your final technical report late your grade will decrease by two thirds of a
step (for example a B to a C+). 

A missing final technical report submission will result in a maximum grade of a D+ at the instructor's discretion.

You can increase your letter grade by *up to* a third of a step (for example a B
to B+) for meeting each of the following conditions (further instructions will
be provided):

-   Complete a weekly submission to the [NYT: What's going on in this
    graph](https://www.nytimes.com/column/whats-going-on-in-this-graph), missing
    no more than two weeks throughout the quarter.
-   Listen to the [99% Invisible podcast on Florence Nightingale: Data Viz Pioneer](https://99percentinvisible.org/episode/florence-nightingale-data-viz-pioneer/). Find the data set to recreate the rose diagram, compare with a stacked bar plot, discuss the story, and argue why Florence needed the rose diagram.
-   Select a data set from
    [TidyTuesday](https://github.com/rfordatascience/tidytuesday) or [Makeover
    Monday](https://www.makeovermonday.co.uk/data/) and use guiding principles
    to create an effective dashboard using Tableau or RShiny to highlight key
    points.
-   Participate in [The Stats+Stories 300th Episode Data Visualization
    Contest](https://community.amstat.org/discussion/the-statsstories-300th-episode-data-visualization-contest#bmc1655f4c-049c-46e9-9553-d30b30d151e4).
-   [Make your own
    website](https://earobinson95.github.io/stat365-calpoly/98-A-options/make-your-own-website.html)
    using Quarto and GitHub to showcase your statistical projects.
-   Complete a technical report project that goes above and beyond the
    requirements. Your job is to convince me in an additional one page document
    why you think your contribution goes the extra mile.

For example, to earn an A in this class you must meet the conditions for a B and
complete three of the "boosting" conditions above to "satisfactory" quality.

Any deviation from the grading policies outlined above will only be to your
benefit.

## Grade Breakdown

1.  **Readings, readiness quizzes, and participation:**

You will be expected to complete assigned readings before coming to class and
participate in class discussions and activities. Additionally, you will be
required to take readiness quizzes to assess your understanding of the material.
Your peer review participation and engagement in class will be evaluated
throughout the quarter.

2.  **Writing prompts and activities:**

Throughout the course, you will be given writing prompts and activities to help
you develop your technical writing skills. These will include assignments such
as summarizing statistical findings, interpreting graphs and tables, and writing
up executive memos. These assignments will be designed to help you practice
communicating complex statistical concepts and analyses clearly and effectively.

3.  **Mini-projects:**

There will be three mini-projects throughout the quarter, each designed to help
you develop and apply your statistical communication skills. The mini-projects
will include:

-   *One Number Story:* In this assignment, you will choose a single number that
    summarizes a key finding from the provided data set and craft a compelling
    story around it, using appropriate statistical evidence to support your
    claims.

-   *Copy the Masters:* For this assignment, you will reproduce graphics from
    FiveThirtyEight. You will use good data visualization principles and
    practicies.

-   *Build your resume:* In this assignment, you will create a resume using
    Overleaf to highlight your education, statistical communication and
    technical skills, statistics and data science projects, and other relevant
    experiences.

Detailed rubrics will be provided for each mini-project to help guide your work
and ensure that you meet the required standards.

4.  **Technical report project:**

The technical report project is the culminating assignment of the course, where
you will have the opportunity to apply all the skills and knowledge you have
acquired throughout the quarter. This will be completed individually. You will be required to choose a research
question, collect and analyze data using appropriate statistical methods, and
produce a technical report that effectively communicates your findings. You will
be provided with guidelines and resources to support your work, and you will
receive feedback and guidance throughout the project. The technical report
project will be evaluated based on its clarity, thoroughness, and adherence to
standard statistical practices.

## Late work

I understand that unexpected events may arise during the quarter that may
prevent you from submitting your assignments on time. To accommodate for this,
you will be given four 24-hour late submission "tokens" to use throughout the
quarter. To use a token, you must (1) send me an email and (2) fill out the deadline extension form [here](https://forms.gle/3uVhdWCN4DN373XE7).

These tokens cannot  be used on your final technical report submission.

Your remaining tokens will be tracked in a Canvas assignment as four points,
with each token worth one point.

# Class Schedule

```{r calendar}
#| echo: false
#| eval: true
#| warning: false
#| fig-width: 8
#| fig-height: 3.5
#| fig-pos: 'h'
class_cal
```

This schedule is **very** tentative and subject to change.

+----------------+------------------------+----------------------------------+
| Week           | Topics                 | Readings                         |
+================+========================+==================================+
| 1: 4/3, 4/5    | Writing about Data     | CwD (ch. 1, ch. 10.3), How to    |
|                |                        | Lie with Stats (ch. 10)          |
|                | Designing a Study      |                                  |
+----------------+------------------------+----------------------------------+
| 2: 4/10, 4/12  | Data Journalism        | CwD (ch. 2, 7), Numbers in the   |
|                |                        | Newsroom (ch. 1, 2)              |
|                | Spreadsheets,          |                                  |
|                | summaries, and pivot   |                                  |
|                | tables                 |                                  |
+----------------+------------------------+----------------------------------+
| 3: 4/17, 4/19  | Descriptive            | CwD (3.1 - 3.2)                  |
|                | Statistics + Tables    |                                  |
|                |                        |                                  |
|                | Reproducibility        |                                  |
+----------------+------------------------+----------------------------------+
| 4: 4/24, 4/26  | Intro to viz           | CwD (4, 8, 9), The Functional    |
|                |                        | Art (intro), Visualize this      |
|                | Color                  | (intro, ch. 1), Storytelling     |
|                |                        | with Data (ch. 7 - 10)           |
|                |                        |                                  |
|                |                        | Design Elements (ch. 3),         |
|                |                        | Envisioning Infromation (ch. 5)  |
+----------------+------------------------+----------------------------------+
| 5: 5/1, 5/3    | Perception +           | Cleveland (ch. 4), Show me the   |
|                | Principles             | numbers (ch. 5) Data Points (ch. |
|                |                        | 5)                               |
|                | Simplification         |                                  |
|                |                        | CwD (ch. 5), Show me the numbers |
|                |                        | (ch. 8), Visual Display (ch. 4,  |
|                |                        | 6)                               |
+----------------+------------------------+----------------------------------+
| 6: 5/8, 5/10   | Statistical Tendencies | Twenty-five Analogies for        |
|                |                        | Explaining Statistical Concepts  |
|                | Writing reports        | (Behar, Grima, and               |
|                |                        | Margo-Almagro)                   |
+----------------+------------------------+----------------------------------+
| 7: 5/15, 5/17  | ADEPT with Professor   |                                  |
|                | Smith                  |                                  |
|                |                        |                                  |
|                | Infographics &         |                                  |
|                | Dashboards             |                                  |
+----------------+------------------------+----------------------------------+
| 8: 5/22, 5/24  | Visit from Career      |                                  |
|                | Center                 |                                  |
|                |                        |                                  |
|                | Finding Internships +  |                                  |
|                | Online Presence        |                                  |
+----------------+------------------------+----------------------------------+
| 9: 5/29, 5/30  | **No Class -** watch   |                                  |
|                | Canvas for             |                                  |
|                | asynchronous material  |                                  |
|                |                        |                                  |
|                | Memorial Day,          |                                  |
|                | Instructor traveling   |                                  |
|                | for an academic        |                                  |
|                | conference             |                                  |
+----------------+------------------------+----------------------------------+
| 10: 6/5, 6/7   | Project Workdays       |                                  |
+----------------+------------------------+----------------------------------+
| Finals         |                        |                                  |
+----------------+------------------------+----------------------------------+

# Course Expectations

You will get out of this course what you put in. The following excerpt was taken
from Rob Jenkins' article "Defining the Relationship" which was published in The
Chronicle of Higher Education (August 8, 2016). This accurately summarizes what
I expect of you in my classroom (and also what you should expect of me).

*"I'd like to be your partner. More than anything, I'd like for us to form a
mutually beneficial alliance in this endeavor we call education.*

*I pledge to do my part. I will:*

-   *Stay abreast of the latest ideas in my field.*
-   *Teach you what I believe you need to know; with all the enthusiasm I
    possess.*
-   *Invite your comments and questions and respond constructively.*
-   *Make myself available to you outside of class (within reason).*
-   *Evaluate your work carefully and return it promptly with feedback.*
-   *Be as fair, respectful, and understanding as I can humanly be.*
-   *If you need help beyond the scope of this course, I will do my best to
    provide it or see that you get it.*

*In return, I expect you to:*

-   *Show up for class each day or let me know (preferably in advance) if you
    have some good reason to be absent.*
-   *Do your reading and other assignments outside of class and be prepared for
    each class meeting.*
-   *Focus during class on the work we're doing and not on extraneous matters
    (like whoever or whatever is on your phone at the moment).*
-   *Participate in class discussions.*
-   *Be respectful of your fellow students and their points of view.*
-   *In short, I expect you to devote as much effort to learning as I devote to
    teaching.*

*What you get out of this relationship is that you'll be better equipped to
succeed in this and other college courses, work-related assignments, and life in
general. What I get is a great deal of professional and personal satisfaction.
Because I do really like you \[all\] and want the best for you."*

# Learning Environment and Support

I am committed to creating a safe and inclusive learning environment where all
students feel respected and supported. If there are any ways I can improve the
classroom environment to make it more welcoming for you, please don't hesitate
to let me know.

If you have a disability and require accommodations to fully participate in the
course, please contact me as soon as possible to discuss how I can best support
you. I also encourage you to register with Cal Poly's Disability Resource Center
(Building 124, Room 119 or at 805-756-1395) to explore additional accommodations
that may be available to you.

If you are experiencing food insecurity, housing instability, or other
challenges that may impact your ability to succeed in this course, please refer
to the resources listed on Canvas under "Student Support Services at Cal Poly."
These resources provide a range of essential support services, including
emergency financial assistance, counseling, and academic support.

I am committed to working with you to ensure that you have the resources and
support you need to succeed in this course. Let's work together to create a
positive and inclusive learning environment for all students.

# Academic Integrity

Academic integrity is a fundamental value of this course and of the university.
Simply put, I will not tolerate cheating, plagiarism, or any other form of
academic dishonesty.

Any incident of academic misconduct, including dishonesty, copying, or
plagiarism, will be reported to the Office of Student Rights and
Responsibilities.

Cheating or plagiarism will result in an incomplete grade for the assignment and
an overall grade deduction of one-third (e.g., B to a B-). In cases of flagrant
cheating, a grade of F for the course may be assigned.

It is important to note that paraphrasing or quoting another's work without
proper citation is a form of academic misconduct. This includes using Chat GPT,
which should only be used to generate ideas and not as a substitute for your own
work.

To ensure academic integrity, please be sure to cite all sources appropriately
and only use Chat GPT in an ethical manner. For more information on academic
misconduct and what constitutes cheating and plagiarism, please see
[academicprograms.calpoly.edu/content/academicpolicies/Cheating](https://academicprograms.calpoly.edu/content/academicpolicies/Cheating).

# Acknowledgments

I would like to acknowledge the contributions of various individuals whose work has been incorporated into this course. Special thanks to Beth Chance, Allan Rossman, Amelia McNamara, and Sara Stoudt for their valuable materials and inspiration.
