---
title: "Stat 365: Statistical Communication"
# author: ""
# date: "January 9, 2023"

format: 
  html:
    self-contained: true
    toc: true
  pdf:
    keep-tex: false
    # include-in-header: "in-header.tex"
    include-in-header:
      text: |
        \makeatletter
        \def\@maketitle{%
          \begin{center}%
          \let \footnote \thanks
            {\LARGE \@title \par}%
            {\large \@author \par}%
            {\large \@date}
          \end{center}%
          \par
          \vskip 1em}
        \makeatother
        \RedeclareSectionCommand[beforeskip=1ex plus -.2ex minus -.2ex,afterskip=.25ex plus -.1ex minus -.1ex]{section}
        \RedeclareSectionCommand[beforeskip=1ex plus -.2ex minus -.2ex,afterskip=.25ex plus -.1ex minus -.1ex]{subsection}
        \RedeclareSectionCommand[beforeskip=1ex plus -.2ex minus -.2ex,afterskip=.25ex plus -.1ex minus -.1ex]{subsubsection}
        \raggedbottom
        \usepackage{enumitem}
        \setlist{nolistsep}
    latex-engine: xelatex

dev: png
geometry: margin=1in

fontfamily: mathpazo
fontsize: 11pt
---

## Contact

**Dr. Emily Robinson**

-	Email: [erobin17@calpoly.edu](mailto:erobin17@calpoly.edu?subject=Stat%20331/531)
-	Office: Building 25 Office 103 (by Statistics Department Office) 

## Course Info

**Class Meeting Times:** 

Mondays/Wednesdays

- Section 01: 1:10pm - 2:00pm
- Section 02: 2:10pm - 3:00pm

**Room:** 33-457 (Math & Science)

## Office Hours

| Day       | Time                                                                           |
|-----------|--------------------------------------------------------------------------------|
| Tuesdays  | 10:10am - 11:00am, in-person (25-103)                                          |
| Thursdays | 10:10am - 11:00am, in-person (25-103)                                          |

```{r}
#| include: false

# knitr::opts_chunk$set(cache=FALSE, dev='pdf')
# knitr::opts_chunk$set(cache=F,
#                       fig.path = 'figs/',
#                       cache.path='cache/',
#                       fig.align = 'center',
#                       echo = F,
#                       warning=F,
#                       message=F)
# knitr::opts_chunk$set(
#                   fig.process = function(x) {
#                       x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
#                       if (file.rename(x, x2)) x2 else x
#                       }
#                   )

library(tidyverse)
library(lubridate)

# Create a calendar for your syllabus ----
# Source: http://svmiller.com/blog/2020/08/a-ggplot-calendar-for-your-semester/
# 1) what is the first Monday of the semester?
# Any number of ways to identify dates in R, but we'll use {lubridate} and the ymd() function here.
# Format: YYYYMMDD. In this example, 4 January 2022.
# What are the full dates of the semester?
semester_dates <- seq(ymd(20230402), ymd(20230617), by=1)
# Weekday(s) of class
class_wdays <- c("Mon", "Wed")

not_here_dates <- c(
  ymd("20230529")
  # Break
  # seq(ymd(20221017), ymd(20221018), by = 1)
  )

exam_dates <- c()

# project_dates <- c(ymd(20221104), ymd(20221205), ymd(20221209))
finals_week <- seq(ymd(20230611), ymd(20230617), by = 1)

# Custom function for treating the first day of the month as the first week 
# of the month up until the first Sunday 
# (unless Sunday was the start of the month)
wom <- function(date) {
    first <- wday(as.Date(paste(year(date), month(date), 1, sep = "-")))
    return((mday(date) + (first - 2)) %/% 7 + 1)
  }
# Create a data frame of dates, assign to Cal
Cal <- tibble(date = seq(ymd(20230401), ymd(20230620), by=1))  %>%
  mutate(mon = lubridate::month(date, label=T, abbr=F),
         wkdy = weekdays(date, abbreviate=T),
         wkdy = fct_relevel(wkdy, "Sun", "Mon", "Tue", "Wed", "Thu","Fri","Sat"), # make sure Sunday comes first
         semester = date %in% semester_dates, 
         # project = date %in% project_dates,
         exams = date %in% exam_dates, # is it an exam?
         not_here = date %in% not_here_dates, # is it a day off?
         exam_wk = date %in% finals_week,
         day = lubridate::mday(date), 
         week = wom(date))
Cal <- Cal %>%
  mutate(category = case_when(
    # project ~ "Project",
    exams ~ "Exam",
    not_here ~ "Cal Poly Holiday",
    semester & wkdy %in% class_wdays & !not_here & !exam_wk | date == "2023-01-17" ~ "Class Day",
    semester ~ "Finals",
    TRUE ~ "NA"
  )) 
class_cal <- Cal %>% 
  ggplot(.,aes(wkdy, week)) +
  theme_bw() +
  theme(aspect.ratio = 1, 
        panel.grid.major.x = element_blank(),
        legend.position = "right",
        # legend.position = c(1, 0), 
        # legend.justification = c(1,0),
        # legend.direction = "vertical", 
        legend.title = element_blank(),
        axis.title.y = element_blank(), 
        axis.title.x = element_blank(),
        axis.ticks.y = element_blank(), 
        axis.text.y = element_blank()) +
  geom_tile(alpha=0.8, aes(fill=category), color="black", size=.45) +
  facet_wrap(~mon, scales = "free", ncol=3) +
  geom_text(aes(label = day, color = semester & (!not_here))) +
  # put your y-axis down, flip it, and reverse it
  scale_y_reverse(breaks=NULL) +
  # manually fill scale colors to something you like...
  scale_color_manual(values = c("FALSE" = "grey80", "TRUE" = "black"), guide = "none") + 
  scale_fill_manual(values=c("Class Day"="purple", 
                             "Finals"="white",
                             "Cal Poly Holiday" = "grey10",
                             "NA" = "white", # I like these whited out...
                             "Exam"="orange"),
                    #... but also suppress a label for a non-class semester day
                    breaks=c("Finals", "Cal Poly Holiday", "Class Day","Exam"))
class_cal

# project_days <- filter(Cal, category == "Project") %>%
#   mutate(topic = c("Project Proposal Due", "Project Presentation Due", "Project Report Due"),
#          time = c("8pm", "8pm", "8pm")) %>%
#   mutate(Date = format(date, "%b %e"),
#          sem_week = week(date))

class_days <- filter(Cal, category %in% c("Class Day")) %>%
  mutate(sem_week = week(date)) %>%
  group_by(sem_week) %>%
  # group_by(sem_week) %>%
  summarize(Date = paste(format(date, "%b %e"), sep = ",", collapse = ", ")) %>%
  mutate(topic = c(
    "",
    "",
    "",
    "",
    "",
    "",
    "",
    "",
    "",
    "")) %>%
  # bind_rows(project_days) %>%
  # bind_rows(exam_days) %>%
  arrange(sem_week) %>%
  select(Date, Topic = topic)
```

# Course Description

Stat 335/531 provides you with an introduction to programming for data and statistical analysis. The course covers basic programming concepts necessary for statistics, good computing practice, and use of built-in functions to complete basic statistical analyses.

# Learning Objectives

This course will teach you the foundations of statistical computing principles in the language of `R`.


# Course Resources

**Textbook**



**Equipment**


# Class Schedule & Topic Outline

This schedule is tentative and subject to change.

```{r calendar}
#| echo: false
#| eval: true
#| warning: false
#| fig-width: 8
#| fig-height: 3.5
#| fig-pos: 'h'
#| fig-cap: "Note: Tuesday, January 17th will follow a Monday class schedule."
class_cal
```

```{r schedule}
#| echo: false
#| eval: true
#| fig-pos: 'h'
class_days %>%
kableExtra::kable(caption = "Tentative schedule of class topics and important due dates", format = "simple")
```

# Course Policies

## Assessment/Grading

Your grade in STAT 331/531 will contain the following components:

Lower bounds for grade cutoffs are shown in the following table. I will not "round up" grades at the end of the quarter. See this `r fontawesome::fa(name = "twitter", fill = "steelblue")` [Twitter thread](https://twitter.com/drseanmullen/status/1604212304622518272?s=46&t=II3oNLTSSrljVPqptoe21g) for advice on *"Playing the lines. Don't be there."*


### General Evaluation Criteria

In every assignment, discussion, and written component of this class, you are
expected to demonstrate that you are intellectually engaging with the material.
I will evaluate you based on this engagement, which means that technically
correct but low effort answers which do not demonstrate engagement or
understanding will receive no credit.

When you answer questions in this class, your goal is to show that you either
understand the material or are actively engaging with it. If you did not achieve
this goal, then your answer is incomplete, regardless of whether or not it is
technically correct. This is not to encourage you to add unnecessary complexity
to your answer - simple, elegant solutions are always preferable to unwieldly,
complex solutions that accomplish the same task.

While this is not an English class, grammar and spelling are important, as is
your ability to communicate technical information in writing; both of these
criteria will be used in addition to assignment-specific rubrics to evaluate
your work.

## Assignment Breakdown


## Attendance & Participation

I do not take formal attendance in this class. However, it is my expectation
that you remain in class and on task until you have finished all your activities
and assignments. Consistent, repeated failure to attend class or actively
participate in portions of the course will affect the demonstration of your
engagement with the course.

If you are feeling ill, please **do not come to class**. Instead, email me,
review the material and work on the participation activity and weekly lab
assignment; then schedule an appointment with me to meet virtually.

## Late Policy


# Course Expectations

You will get out of this course what you put in. The following excerpt was taken
from Rob Jenkins’ article “Defining the Relationship” which was published in The
Chronicle of Higher Education (August 8, 2016). This accurately summarizes what
I expect of you in my classroom (and also what you should expect of me).

*“I’d like to be your partner. More than anything, I’d like for us to form a mutually beneficial alliance in this endeavor we call education.*

*I pledge to do my part. I will:*

+ *Stay abreast of the latest ideas in my field.* 
+ *Teach you what I believe you need to know; with all the enthusiasm I possess.*
+ *Invite your comments and questions and respond constructively.*
+ *Make myself available to you outside of class (within reason).*
+ *Evaluate your work carefully and return it promptly with feedback.* 
+ *Be as fair, respectful, and understanding as I can humanly be.*
+ *If you need help beyond the scope of this course, I will do my best to provide it or see that you get it.* 

*In return, I expect you to:*

+ *Show up for class each day or let me know (preferably in advance) if you have some good reason to be absent.* 
+ *Do your reading and other assignments outside of class and be prepared for each class meeting.*
+ *Focus during class on the work we’re doing and not on extraneous matters (like whoever or whatever is on your phone at the moment).* 
+ *Participate in class discussions.*
+ *Be respectful of your fellow students and their points of view.* 
+ *In short, I expect you to devote as much effort to learning as I devote to teaching.*

*What you get out of this relationship is that you’ll be better equipped to succeed in this and other college courses, work-related assignments, and life in general. What I get is a great deal of professional and personal satisfaction. Because I do really like you [all] and want the best for you.”*

### Be Creative!

# University Policies

See [academicprograms.calpoly.edu/content/academicpolicies](https://academicprograms.calpoly.edu/content/academicpolicies/index).

## Learning Environment and Support

I believe everyone is capable of learning statistics and programming with proper support. It is my goal for everyone to feel safe and comfortable in my classroom. If
there is any way I can make the course more welcoming for you, please do not
hesitate to ask.

In particular, if you have a disability, I will gladly work with you to make
this class accessible.

I encourage you to also contact the Disability Resource Center (Building 124,
Room 119 or at 805-756-1395), who can help you register for extra accommodations
such as extended exam time.

If you are having difficulty affording groceries, lacking a safe & stable place to live, or needing additional essential supports, please see Canvas for a list of Student Support Services at Cal Poly.

## Academic Integrity and Class Conduct

Simply put, I will not tolerate cheating or plagiarism.

Any incident of dishonesty, copying, exam cheating, or plagiarism will be
reported to the Office of Student Rights and Responsibilities.

Cheating will earn you a grade of 0 on the assignment and an overall grade
penalty of at least 10%. In circumstances of flagrant cheating, you may be given
a grade of F in the course.

Paraphrasing or quoting another’s work without citing the source is a form of
academic misconduct. This includes the `R` code produced by someone else!
Writing code is like writing a paper, it is obvious if you copied-and-pasted a
sentence from someone else into your paper because the way each person writes is
different.

Even inadvertent or unintentional misuse or appropriation of another’s work
(such as relying heavily on source material that is not expressly acknowledged)
is considered plagiarism. If you are struggling with writing the `R` code for an
assignment, please reach out to me. I would prefer that I get to help you rather
than you spending hours Googling things and get nowhere!

If you have any questions about using and citing sources, you are expected to
ask for clarification.

For more information about what constitutes cheating and plagiarism, please see
[academicprograms.calpoly.edu/content/academicpolicies/Cheating](https://academicprograms.calpoly.edu/content/academicpolicies/Cheating).

